---
title: "Functional Censored Quantile Regression"
author: "Fei Jiang and Qing Cheng"
output:
  pdf_document:
    extra_dependencies: ["float"]
    number_sections: yes
---

```{r setup,warning = FALSE, include=FALSE}
knitr::opts_chunk$set( echo = TRUE)
```
# Introduction
This vignette provides an introduction to the `FCQR` package. R package `FCQR` implements method for Functional Censored Quantile Regression. YouInstall The development version of FCQR can be installed by use of the following command in R.
```
remotes::install_github("QingCheng0218/FCQR@main");
```

Load the package using the following command:
```{r}
library(FCQR);
```

# Simulation Study

## Finite Sample Performance Without IGACV Procedure
In this section, we conduct simulation studies to assess the finite-sample performance of the FCQR procedure. We first load the dependent  R packages as follows:
```{r, message=F, warning=F}
library(xtable); library(survival); library(fda);
library(quantreg); library(ggplot2); library(MASS);
```
We generate the event time from the model,
\begin{eqnarray*}
\log(T_i) = b_1 X_{1i} + \int_0^1 Z_i(s)\phi(s) ds + \left\{
b_2 X_{2i} +  \int_0^1 Z_i(s) ds \right\} \epsilon_i,
\end{eqnarray*}
where $X_{1i}$ is a normal variate with variance $\sigma_{z}^2$, $X_{2i}$ is a variate uniformly distribution on $[0, c_0]$, and $\epsilon_i$ is a normal random error with mean zero and variance $\sigma_{\epsilon}^2$. The detailed information of functional covariate $Z_i(s)$ and $\phi(s)$ can be found in Section 5.1 of our paper.
We set $\sigma_{\epsilon} = 0.2$, $\nu = 2.5$, $n = 200$ and $\lambda_0$ is used to control the censored rate.
```{r}
n <- 200; v <- 1; sdx <- 0.2; sdz <- 1; lam0 <- 5;

t <- seq(0, 1, 0.01);  m <- length(t); deltat <- (diff(t)[1]);
nk <- 50; c0 <- 1;
b0 <- c(1, 2); lb0 = length(b0);

grd <- seq(0.1, 0.9, 0.1/5);
qv <- c(which(eleq(0.3, grd)), which(eleq(0.4, grd)), which(eleq(0.5, grd)), 
        which(eleq(0.6, grd)), which(eleq(0.7, grd)));
```
We using the following function to B-spline bases.
```{r}
# ---------------------------------------------------------#
# creat B-spline basis.
nbasis <- 5; norder <- 4;
basis <- create.bspline.basis(range(t),nbasis = nbasis, norder = norder)
nbasis <- basis$nbasis
valueBasis <-eval.basis(t, basis)
# ---------------------------------------------------------#
```

```{r}

nsim = 10;
Base0.m <- Sigb0.m <- array(0, dim = c(nsim, length(qv), lb0));
Bfun.m <- Sigfun.m <- array(0, dim = c(nsim, m, length(qv)))
cdata <- matrix(0, ncol = nsim, nrow = n);
itr <- 0; tt <- 1;
repeat{
  itr <- itr +1
  # ---------------------------------------------------------#
  # Generate Sim function.
  temp <- simfun(t, v, n, sdx, sdz, b0,  lam0, c0);
  Zb <- temp$Zb # baseline covariate.
  time <- temp$Timedata # survival time and censoring time.
  Zs <- temp$mZ # functional covariate.
  cdata[, itr] <- 1*(time[, 1] <= time[, 2])
  # ---------------------------------------------------------#
  res.itr <- try(Fcqr(time, Zs, Zb, t, grd, qv, valueBasis));
  if(class(res.itr) != 'try-error'){
    Base0.m[itr, , 1:lb0] <- res.itr$base0;
    Sigb0.m[itr, ,1:lb0] <- res.itr$sigb0;
    Bfun.m[itr, , ] <- t(res.itr$Bfun);
    Sigfun.m[itr, , ] <- t(res.itr$sigfun);
    tt <- tt + 1;
  }
  if(tt%%100==0) {print(tt);print(itr);print(Sys.time())}
  if(tt == nsim + 1){
    break
  }
}

cenRate = 1 - mean(colMeans(cdata))
```

```{r, message=F, warning=F}
#**********************************#
# The result of coefficient function.
#**********************************#
alfun <- simBX(t, v, nk)$B0
plot.m <- alfun.be(alfun, Sigfun.m, Bfun.m, grd, qv, sdx);
tau.id <- 3;
resplot6 <- alphaplot.be(t, grd, qv, tau.id, plot.m, -10, 1);
```


```{r}
#**********************************#
# The result of Beta0
#**********************************#
lgrd <- length(grd);
btautrue <- cbind(rep(b0[1], lgrd), b0[1]*qnorm(grd, 0, sdx))
b0res <- rbind(b0res.be(Base0.m[,,1], Sigb0.m[,,1], btautrue[,1], qv),
               b0res.be(Base0.m[,,2], Sigb0.m[,,2], btautrue[,2], qv));
colnames(b0res) = c("BIAS", "SD", "SE", "CP");
rownames(b0res) = paste0(rep(c("$\\hat\\beta_1(","$\\hat\\beta_2("), 5),
                         rep(grd[qv], each  = 2),rep(")$", 10));
```

Table 1 shows the estimation result for $\beta(\tau)$, the result of time-varying estimator can be found in Figure 1. One can increase the number of replications, denoted by \textit{nsim}, to obtain more accurate results.

```{r, echo=FALSE, results='asis', fig.pos="H"}
knitr::kable(b0res, caption = "\\label{testPlot}Estimation results for $\\hat \\beta(\\tau) = (\\hat \\beta_1 (\\tau_1), \\hat \\beta_2(\\tau_2)$ at different quantiles with a censoring rate of 20%.")
```

```{r gg-oz-gapminder,warning = FALSE, fig.width = 6, fig.height=4,   fig.cap = "\\label{testPlot}Estimated coefficient function $B^T_{rS}(s)\\hat \\gamma_{S}(0.5)$  under a censoring rate of 20%.", fig.pos="H"}
resplot6
```

## Performance of IGACV for Knot Selection
In this section, we mimic the real data to illustrate the performance of the IGACV procedure. 
```{r, message=F, warning=F}
rm(list = ls())
library(FCQR);
library(xtable); library(survival); library(fda);
library(quantreg); library(ggplot2); library(MASS);
```
In particular, we generate $n = 297$ survival time from the model
\begin{eqnarray*}
\log(T_i) = \int_0^1 Z_i(s)\phi(s) ds + \left\{
b_1 X_{2i} +  \int_0^1 Z_i(s) ds \right\} \epsilon_i,
\end{eqnarray*}
where $Z_i(s) = SBP_i(s) + U_i(s)$ and $Z_{1i} = \text{min} (SBP_i) +U_{1i}$ with $SBP_i(s)$ corresponding to the systolic blood pressure(SBP) trajectory for the $i-$th individual, and $U_i(s)$ and $U_{1i}$ are uniform random variables on $[-c_0, c_0]$.
```{r}
# ------------------------------------------------------------
t <- time; zsm <- MSdata[, 3:98]; 
c0 <- 10;  sdx <- 0.2; sdz <- 1; beta1 <- 0.1; norder <- 4;
# ------------------------------------------------------------
grd <- seq(0.1, 0.9, 0.005/5);
qv <- c(which(eleq(0.3, grd)), which(eleq(0.4, grd)), which(eleq(0.5, grd)), 
        which(eleq(0.6, grd)), which(eleq(0.7, grd)));
# ------------------------------------------------------------
```
We evaluate IGACV and IBIC over the choices of $d_s = 4, 5, 6, 7, 8$.
```{r}
bsnb <- c(4, 5, 6, 7, 8);
nsim <- 10; lam0 <- 1 # for 20% censoring rate.
resop = optbasis(nsim, bsnb, zsm, lam0, c0, beta1, sdz, sdx,
                 grd, t, norder, qv);
resoptb = resop$inopt;
rownames(resoptb) = c("IGACV", "IBIC")
colnames(resoptb) = bsnb;
```

```{r, echo=FALSE, results='asis', fig.pos="H"}
knitr::kable(resoptb, caption = "\\label{testPlot}Optimal $d_S$ based on the IGACV and IBIC criteria under different censoring rate of 20% over 10 simulations.")
```

IGACV chooses $d_S = 6$ as the optimal number of basis functions, see Table 2. Next, we present the estimation results for $d_S$ in Table 3 and Figure 2.
```{r, message=F, warning=F}
# -------------------------------------------------------------
nbasis <- 6;
btautrue <- beta1*qnorm(grd, 0, sdx);
# -------------------------------------------------------------
# report tau equals 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8;
grd <- seq(0.1, 0.9, 0.1/5);
qv <- c(which(eleq(0.2, grd)), 
        which(eleq(0.3, grd)), which(eleq(0.4, grd)),
        which(eleq(0.5, grd)), which(eleq(0.6, grd)), 
        which(eleq(0.7, grd)), which(eleq(0.8, grd)));
# -------------------------------------------------------------
SIMRES <- Mimicfun(zsm, lam0, c0, beta1, sdz, sdx, nsim,
                   grd, t, nbasis, norder, qv);

Base0.m <- SIMRES$Base0.m;
Sigb0.m <- SIMRES$Sigb0.m;
Sigfun.m <- SIMRES$Sigfun.m;
Bfun.m <- SIMRES$Bfun.m;

b0res <- b0res.be(Base0.m, Sigb0.m, btautrue, qv)
colnames(b0res) = c("BIAS", "SD", "SE", "CP");
rownames(b0res) = seq(0.2, 0.8, 0.1);

n <- nrow(MSdata);
alfun <- SimMimicfun(1, zsm, t, n, sdz, sdx, c0, beta1, lam0)$alp;
plotres <- alfun.be(alfun, Sigfun.m, Bfun.m, grd, qv, sdx);
tau.id <- 4;
mimicplot <- alphaplot.be(t, grd, qv, tau.id, plotres, -16, 16);
```

```{r, echo=FALSE, results='asis', fig.pos="H"}
knitr::kable(b0res, caption = "\\label{testPlot}Estimation results for $\\hat \\beta(\\tau)$ at different quantiles with a censoring rate of 20%.")
```

```{r gg-oz-gapminder1,warning = FALSE, fig.width = 6, fig.height=4,   fig.cap = "\\label{testPlot}Estimated coefficient function $B^T_{rS}(s)\\hat \\gamma_{S}(0.5)$ with $d_S = 6$ under a censoring rate of 20%.", fig.pos="H"}
mimicplot;
```

Again, the accuracy of results can be improved by increasing the number of simulations.

# Stroke Application
In this section, we identify functional relationship between ambulatory blood pressure trajectories and clinical outcomes in stroke patients.
We first apply the IGACV and IBIC criteria to select $d_S$.
```{r, message=F, warning=F}
rm(list=ls());
library(FCQR);
# --------------------------------------------------
library(imputeTS); library(survival); library(survminer); library(fda);
library(quantreg); library(statmod); library(xtable); library(Cairo);
# --------------------------------------------------

a0 <- 0; b0 <- 1; gn <- 20;
norder <- 4;
grd <- seq(0.1, 0.9, 0.005/5); lgrd <- length(grd);
qv <- c(which(eleq(0.2, grd)),
        which(eleq(0.3, grd)), which(eleq(0.4, grd)), which(eleq(0.5, grd)), 
        which(eleq(0.6, grd)), which(eleq(0.7, grd)), which(eleq(0.8, grd)));
# -----------------------------------------------------#
# using GACV method to determine optimal nbasis
CBA <- c(1, 2, 3, 4, 5)
BICR <- GACVR <- matrix(0, ncol=length(CBA), nrow=lgrd)
NBS <- rep(0,length(CBA))

n <- dim(MSdata)[1]
cn <- (log(n))^(1/2)
for(i in 1:length(CBA)){
  cba <- CBA[i];
  res <- predfun(MSdata, a0, b0, gn, time, norder, cba);
  TM <- res$TM; DEL <- res$DEL;
  nbasis <- res$nbasis; ns <- nbasis + 1;
  pre.y <- res$pre.y;
  bicres <- bicfun(n, ns, cn, TM, pre.y, DEL, grd);
  GACVR[, i] <- bicres$gacv;
  BICR[, i] <- bicres$bic;
  NBS[i] <- nbasis;
}

bicres <- optfun(grd, BICR);
gacvres <- optfun(grd, GACVR);
resoptReal <- rbind(gacvres[2, ], bicres[2, ]);
rownames(resoptReal) <- c("IGACV", "IBIC");
colnames(resoptReal) <- 4:8;
# -----------------------------------------------------#
```

```{r, echo=FALSE, results='asis', fig.pos="H"}
knitr::kable(resoptReal, caption = "\\label{testPlot}Selection of the optimal number of basis functions using the IGACV and IBIC criteria respectively for the stroke data.")
```

As shown in Table 4, the IGACV reaches the minimum value at $d_S = 7$. Based on $d_S = 7$, we estimate $\beta(\tau)$ and $\alpha(s, \tau)$ using the following command.

```{r}
cba <- 4;
covres <- intzfun0(MSdata, a0, b0, gn, time, norder, cba)
covdata <- covres$covdata
Z <- covres$Z
valueBasisT <- covres$valueBasisT
reswei <- weifun(MSdata, covdata)
# weight0 <- reswei
weight0 <- rep(1, nrow(MSdata))
coefres <- coeffun0(valueBasisT, covdata,  MSdata,
                    grd, qv, weight0, Z, time)
# -----------------------------------------------------
# Result of baseline covariate
base0 <- coefres$base0
sigb0 <- coefres$sigb0

c2 <- base0 + 1.96 * sqrt(sigb0);
c1 <- base0 - 1.96 * sqrt(sigb0);

ci0 <- paste("(",round(c1,5)," ,",round(c2,5),")",sep="")
resbase0 <- rbind(round(base0, 5), ci0)
rownames(resbase0) <- c("Estimators", "95% CI ")
resbase = resbase0[, 2:5];
colnames(resbase) = seq(0.3, 0.6, 0.1);
```

```{r, echo=FALSE, results='asis', fig.pos="H"}
knitr::kable(resbase, caption = "\\label{testPlot}Estimates $\\hat \\beta(\\tau)$ and the 95% bootstrap confidence intervals (CIs) at different values of $\\tau$ for the stroke data.")
```

```{r}
# Result of time dependent covariate
Bfun <- coefres$Bfun
sigfun <- coefres$sigfun
DA00 <- DA01 <- DA02 <- matrix(0, ncol = length(time), nrow = length(grd[qv]))
for(j in 1:length(grd[qv])){
  DA00 <- Bfun;
  bt <- DA00[j, ];
  vc <- sigfun[j, ];
  DA01[j, ] <- bt + 1.96 * sqrt(vc);
  DA02[j, ] <- bt - 1.96 * sqrt(vc);
}

tau.id <- 3;
bt <- Bfun[tau.id, ];
vc <- sigfun[tau.id, ];
da00 <- bt;
da01 <- bt + 1.96 * sqrt(vc);
da02 <- bt - 1.96 * sqrt(vc)

yid <- 0.5; max.y <- yid; min.y <- -yid;
t <- seq(19.25, 43, length.out = 96);
realplot = bandplot(da00, da01, da02, tau.id, t, min.y, max.y);
```

```{r real1,warning = FALSE, fig.width = 5, fig.height=4,  fig.cap = "\\label{testPlot}Estimated coefficient function $B^T_{rS}(s)\\hat \\gamma_{S}(\\tau)$ with $d_S = 7$.", fig.pos="H"}
realplot;
```

Table 5 shows that there are significant negative effects of $Z_{1i}$ at all considered quantiles, which suggests that hypertension is an important risk factor for stroke. Figure 3 displays the estimated $\alpha(s, \tau)$ and the corresponding 95% pointwise bootstrap confidence intervals with $\tau = 0.4$.